Agile Formality: A Mole of Software Verification Practices

### 2010-07-02 papers/ModelBasedTestCaseGenerationDataStreamProtocol/paper.tex
new abstract

@@

\ifpdf 
  \usepackage{epstopdf}
  \usepackage[pdftex,bookmarks=false,a4paper=false,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Joseph R. Kiniry and Vieri del Bianco"]{hyperref} 
\else
  \usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

@@

\author{Vieri del Bianco, Joseph R. Kiniry and Dragan Stosic\\
  UCD CASL: Complex and Adaptive Systems Laboratory and\\
  School of Computer Science and Informatics,\\
  University College Dublin,\\
  Belfield, Dublin 4, Ireland,\\
  vieri.delbianco@ucd.ie, kiniry@acm.org and dragan.stosic@gmail.com\\
}

@@

Agile and formal people do not blend nicely. 
These two communities do not talk to or learn from each other. 
The problem-solving approaches and mind attitudes differ widely, but this clash of view points can be exploited into new development processes that actually blend, rather than fight, or simply compose, practices from the two worlds. 
This paper summarizes a complex case study, showing that it is not only possible, but tasty, to combine the ``garlic'' of formal methods and the ``chocolate'' of agile programming: a ``Mole'' of software verification practices.

@@

% previous work: a common approach
As already mentioned, in all these attempts, the hand made test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~/cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.

@@

% how we want to achive it
The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem in an highly iterative development process.
We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how hand made and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices.

>>>>

%
% "Agile Formality: A Mole of Software Verification Practices"
% for Agile Methods+Formal Methods 2010
% Vieri del Bianco, Joseph Kiniry, Dragan Stosic
%

@@

\ifpdf 
  \usepackage{epstopdf}
  \usepackage[pdftex,bookmarks=false,a4paper=true,plainpages=false,naturalnames=true, colorlinks=true,pdfstartview=FitV,linkcolor=blue,citecolor=blue,urlcolor=blue, pdfauthor="Vieri del Bianco and Joseph R. Kiniry and Dragan Stosic"]{hyperref} 
\else
  \usepackage[dvips,linkcolor=blue,citecolor=blue,urlcolor=blue]{hyperref}
\fi

@@

\author{Vieri del Bianco and Dragan Stosic\\
  UCD CASL: Complex and Adaptive Systems Laboratory and\\
  School of Computer Science and Informatics,\\
  University College Dublin, Ireland\\
  \texttt{vieri.delbianco@ucd.ie} and \texttt{dragan.stosic@gmail.com}\\
  \\
  Joseph R.~Kiniry \\
  IT University of Copenhagen, Denmark\\
  \texttt{kiniry@itu.dk}
}

@@

  Members of the agile programming and formal methods communities do
  not always see eye-to-eye.  These two communities often do not talk
  to or learn from each other.  Only recently, as highlighted by the
  September 2009 issue of IEEE Software, the IFIP workshop on
  balancing agility and formalism in software engineering, and the
  first edition of the first international workshop for formal methods
  and agile methods, have ideas from the two communities begun to be
  synthesized.  While the problem-solving approaches and psychological
  attitudes of members of the two communities differ widely, we
  exploit this clash of viewpoints, synthesizing a new development
  processes that actually blends, rather than mashes together, best
  practices from the two worlds.  This paper summarizes our process
  and a supporting complex case study, showing that it is not only
  possible, but tasty, to combine the ``chili pepper'' of formal
  methods and the ``chocolate'' of agile programming, thus producing a
  tasty ``mole'' of software engineering practices.

@@

% previous work: a common approach
As already mentioned, in all these attempts, the hand made test suites are automatically generated using the system formal specification, or completely replaced by theorem provers or static checkers. 
The requisite to enable these approaches is, at least, a specification of the entire system under development to enable simple verification techniques as run time verification (obtained by assertions and design by contract~\cite{Meyer1997} preconditions, postconditions and invariants) and test suites automatic generation~\cite{Cheon2002,Cheon2004,Cheon2005}.
At most, a complete and sound system specification, usually supported by an appropriate formal language (to specify model and properties) and a constrained programming language, to be able to actually prove properties about the system under development~\cite{CatanoHuisman02,DetlefsNelsonSaxe2005,KiniryCok04}.

@@

% how we want to achive it
The main specific problem, by now never addressed, is how to connect development artifacts of the two worlds together: we will define activity cycles, similar to what is found in agile methodologies (especially in XP), to solve this problem in an highly iterative development process.
We will show how tests can drive the formal specification, how the formal specifications can drive tests and code development, how hand made and automated tests can coexist and support each other, how the unspecified parts of the system can be incrementally specified.
That is, we will show how to blend, and not compose, formal and agile verification practices: a Mole of verification practices\footnote{A mole (\emph{'m\=ol\=a}) is a highly spiced Mexican sauce made chiefly from chili peppers and chocolate, served with meat.}
